MATH 3610 - Project \#1

### 2015-09-29 proj1.tex
changed variable names

Suppose there are two groups of people, with $S$ representing the number of 
	sick people, and $H$ represnting the number of healthy people.
Suppose the probability of a single healthy person becoming sick is proportional
	only to the total number of sick people.
Since there are $H$ healthy people, the expected number of healthy people
	who become sick should be proportional to $S H$.
This model does not take into account recovery from sickness.
If we introduce a parameter $r$, we can express this relationship as a system
	of differential equations:

\begin{align*}
\frac{dH}{dt} & = - r S H \\
\frac{dS}{dt} & = r S H \\
\end{align*}

Notice that the total number of people, $S + H$, is conserved, i.e.,

\[ \frac{d}{dt} \left( S + H \right) = \frac{dS}{dt} + \frac{dH}{dt}
	= r S H - r S H = 0 \]

If we introduce another parameter, $K$, to represent the total population,
	such that $S + H = K$ for all time, then the differential equations
	can be decoupled:

\begin{align*}
\frac{dH}{dt} & = - r (K - H) H \\
\frac{dS}{dt} & = r S (K - S) \\
\end{align*}

These are two independant logistic models.

@@

Suppose we introduce a third group, $V$, representing people who are vaccinated	
	against the disease.
Suppose we randomly vaccinate members of the healthy population,
	a constant number each time step.
This introduces a third parameter, $r_v$, the rate of vaccinations.
The vaccinated population isn't seceptible to the disease.
This process is modeled by the following system of diffential equations:

\begin{align*}
\frac{dS}{dt} & = r S H \\
\frac{dH}{dt} & = - r S H - r_v \\
\frac{dV}{dt} & = r_v
\end{align*}

@@

Again, we find the conserved quantity $\left( S + H + V \right)$ and
	name it $K$.
For the model to be valid, we need $H = K - S - V \geq 0$.
Therefore, the model is valid as long as:

\[ K \geq S + r_v t \]

@@

In the above figure, once $S(t)$ hits the diagonal line,
	it stays horizontal and keeps its value.

@@

\begin{align*}
\frac{dS}{dt} & = r S \left( K - S - r_v t \right)\\
\frac{dH}{dt} & = - r H \left( K - H - r_v t \right) - r_v \\
\frac{dV}{dt} & = r_v t
\end{align*}

We present a solution for $S(t)$:

\begin{align}
S(t) & = 
	\frac{e^{ r t (K - \frac{ r_v t}{2})}}
	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t - K ) \right)
			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
\end{align}

From the above expression for the boundary of the valid region,
	we get the following implicit expression for $t_f, S(t_f)$

\begin{align}
S(t_f) = 
	\frac{e^{ r t_f (K - \frac{ r_v t_f }{2})}}
	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t_f - K ) \right)
			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
	& = K - r_v t_f 
\end{align}

Here is a family of values of $S(t)$.
The total population and the rate of distributing vaccines are
	held constant, while each curve corresponds to a different
	value of the growth rate constant.
Each curve starts with an initial condition at 5 percent of the
	population, and the growth rate determines how many people
	overall are infected.

@@

The model above modeled how one population responds to a combination
	of disease spreading and vaccines.
The model depends on $K$, the total population, $S(0)$, the initial
	number of sick people, $r$, the rate of infection growth, 
	and $r_v$, the vaccination rate.
Of these parameters, in a real world setting, we can expect to know $K$
	very precicely from census data, approximate $S(0)$ as small,
	have very little idea about $r$, and control $r_v$.

I want to extend this problem to two independant populations. 
Each population will get different sets of constants,
	$K_1$ and $K_2$, $S_1(0)$ and $S_2(0)$, $r_1$ and $r_2$.
The goal will be to figure out how to optimally allocate the
	vaccines between the two populations.
The decision variable, $d$, will be the fraction of the total vaccination
	rate I allocate to each population.
Population 1 will be vaccinated at a rate $d r_v$, while population
	2 will be vaccinated at a rate $(1 - d) r_v$.
In this model, I try to minimize the final number of sick people in
	each population.
This will give me a pareto front.

@@

>>>>

Suppose there are two groups of people, with $I$ representing the number of 
	infected people, and $S$ represnting the number of people seceptible for the
	disease.
Suppose the probability of a single healthy person becoming sick is proportional
	only to the total number of sick people.
Since there are $S$ healthy people, the expected number of healthy people
	who become sick should be proportional to $S I$.
This model does not take into account recovery from sickness.
If we introduce a parameter $r$, we can express this relationship as a system
	of differential equations:

\begin{align*}
\frac{dI}{dt} & = - r S I \\
\frac{dS}{dt} & = r S I \\
\end{align*}

Notice that the total number of people, $S + I$, is conserved, i.e.,

\[ \frac{d}{dt} \left( S + I \right) = \frac{dS}{dt} + \frac{dI}{dt}
	= r S I - r S I = 0 \]

If we introduce another parameter, $K$, to represent the total population,
	such that $S + I = K$ for all time, then the differential equations
	can be decoupled:

\begin{align*}
\frac{dI}{dt} & = - r (K - I) I \\
\frac{dS}{dt} & = r S (K - S) \\
\end{align*}

The two differential equations shown here are now two independant, uncoupled
	differential equations which we can solve seperately.

@@

Suppose we introduce a third group, $V$ for vaccinate, representing people who are 
	immune from the diesease.
Right now, the only way to become immune is through vaccination,
	but we may add recovery in subsequent models.
Suppose we randomly vaccinate members of the healthy population,
	a constant number each time step.
This introduces a third parameter, $r_v$, the rate of vaccinations.
The vaccinated population isn't seceptible to the disease.
This process is modeled by the following system of diffential equations:

\begin{align*}
\frac{dI}{dt} & = r S I \\
\frac{dS}{dt} & = - r S I - r_v \\
\frac{dV}{dt} & = r_v
\end{align*}

@@

Again, we find the conserved quantity $\left( S + I + V \right)$ and
	name it $K$.
For the model to be valid, we need $S = K - I - V \geq 0$.
Therefore, the model is valid as long as:

\[ K \geq I + r_v t \]

@@

In the above figure, once $I(t)$ hits the diagonal line,
	it stays horizontal and keeps its value.

@@

\begin{align*}
\frac{dI}{dt} & = r I \left( K - I - r_v t \right)\\
\frac{dS}{dt} & = - r S \left( K - S - r_v t \right) - r_v \\
\frac{dV}{dt} & = r_v t
\end{align*}

We present a solution for $I(t)$ \cite{walph}:

\begin{align}
I(t) & = 
	\frac{e^{ r t (K - \frac{ r_v t}{2})}}
	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t - K ) \right)
			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
\end{align}

From the above expression for the boundary of the valid region,
	we get the following implicit expression for $t_f, I(t_f)$

\begin{align}
I(t_f) = 
	\frac{e^{ r t_f (K - \frac{ r_v t_f }{2})}}
	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t_f - K ) \right)
			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
	& = K - r_v t_f 
\end{align}

Here is a family of values of $I(t)$.
The total population and the rate of distributing vaccines are
	held constant, while each curve corresponds to a different
	value of the growth rate constant.
Each curve starts with an initial condition at 5 percent of the
	population, and the growth rate determines how many people
	overall are infected.

@@

The model above modeled how one population responds to a combination
	of disease spreading and vaccines.
The model depends on $K$, the total population, $I(0)$, the initial
	number of sick people, $r$, the rate of infection growth, 
	and $r_v$, the vaccination rate.
Of these parameters, in a real world setting, we can expect to know $K$
	very precicely from census data, approximate $I(0)$ as small,
	have very little idea about $r$, and control $r_v$.

I want to extend this problem to two independant populations. 
Each population will get different sets of constants,
	$K_1$ and $K_2$, $I_1(0)$ and $I_2(0)$, $r_1$ and $r_2$.
The goal will be to figure out how to optimally allocate the
	vaccines between the two populations.
The decision variable, $d$, will be the fraction of the total vaccination
	rate I allocate to each population.
Population 1 will be vaccinated at a rate $d r_v$, while population
	2 will be vaccinated at a rate $(1 - d) r_v$.
In this model, I try to minimize the final number of sick people in
	each population.
This will give me a pareto front.

@@

\begin{thebibliography}{9}
\bibitem{walph}
	Wolfram Research Inc.,
	\emph{WolframAlpha},
	Retrieved September 26, 2015
	http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx
	http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29
\end{thebibliography}
